babette: BEAUti 2, BEAST2 and Tracer for R

### 2018-05-17 article/article.tex
Put manuscript in 'article' folder

  \textbf{1. }
    In the field of phylogenetics, 
    BEAST2 is one of the most widely used software tools. 
    It comes with the graphical user interfaces BEAUti 2, DensiTree and Tracer,
    to create BEAST2 configuration files and to interpret BEAST2's output files. 
    However, when many different alignments or model 
    setups are required, a workflow of graphical user interfaces is cumbersome. \\
  \textbf{2. }
    Here, we present a free, libre and open-source package, \verb;babette;: 
    'BEAUti 2, BEAST2 and Tracer for R', for the R programming language. 
    \verb;babette; creates BEAST2 input files, runs BEAST2 and parses its results, 
    all from an R function call. \\
  \textbf{3. }
    We describe \verb;babette;'s usage and the novel functionality it provides
    compared to the original tools and we give some examples. \\
  \textbf{4. }
    As \verb;babette; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package. \\
\end{abstract}

\begin{abstract}
  \textbf{1. }
    In de fylogenetica is BEAST2 een van de meest gebruikte hulpprogramma's. 
    Het is gebundeld met de grafische gebruiksinterface BEAUti 2, DensiTree en Tracer, 
    om BEAST2-configuratiebestanden te maken en om BEAST2-outputbestanden te interpreteren. 
    Echter, als veel verschillende aligneringen of modelopzetten nodig zijn, 
    is een werkvolgorde van meerdere grafische gebruiksinterfaces onhandig. \\
  \textbf{2. }
    Hier presenteren we een gratis, vrij en open-source package, \verb;babette;: 
    'BEAUti 2, BEAST2 en Tracer voor R', voor de programmeertaal R. 
    \verb;babette; schrijft BEAST2-configuratiebestanden, start BEAST2 and verwerkt de resultaten, 
    alles met een enkele R functie-aanroep. \\
  \textbf{3. }
    We beschrijven hoe \verb;babette; te gebruiken is 
    en de nieuwe mogelijkheden die het biedt vergeleken met de originele programma's, 
    aan de hand van enkele voorbeelden. \\
  \textbf{4. }
    Omdat \verb;babette; ontworpen is voor uitbreidbaarheid en hoge kwaliteit, 
    sluiten we af met het beschrijven van de verdere ontwikkeling van dit package. \\
\end{abstract}

@@

However, BEAUti cannot be called from a command-line script.
This implies that when the user 
wants to explore the consequences of various settings, this must be done manually.
This is the managable workflow when using a few alignments and doing a superficial 
analysis of sensitivity of the reconstructed tree to model settings. 
For exploring many trees (for instance from simulations), 
when doing when doing a sliding-window analysis on a genomic alignment, 
or for a more thorough sensitivity analysis, one would like to loop through 
multiple (simulated or shortened) alignments, nucleotide substitution models, 
clock models and tree priors. 
One such tool to replace BEAUti is \verb;BEASTmasteR; (\cite{beastmaster}),
which focuses on morphological traits and tip-dating, but also 
supports DNA data. \verb;BEASTmasteR;, however, requires hundreds of
lines of R code to setup the BEAST2 model configuration 
and a Microsoft Excel file to specify alignment files.

>>>>

  \textbf{1. }
    In the field of phylogenetics, 
    BEAST2 is one of the most widely used software tools. 
    It comes with the graphical user interfaces BEAUti 2, DensiTree and Tracer,
    to create BEAST2 configuration files and to interpret BEAST2's output files. 
    However, when many different alignments or model 
    setups are required, a workflow of graphical user interfaces is cumbersome. \\
  \textbf{2. }
    Here, we present a free, libre and open-source package, \verb;babette;: 
    'BEAUti 2, BEAST2 and Tracer for R', for the R programming language. 
    \verb;babette; creates BEAST2 input files, runs BEAST2 and parses its results, 
    all from an R function call. \\
  \textbf{3. }
    We describe \verb;babette;'s usage and the novel functionality it provides
    compared to the original tools and we give some examples. \\
  \textbf{4. }
    As \verb;babette; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package. \\

  \begin{center}
  \textbf{Samenvatting}
  \end{center}
  \textbf{1. }
    In de fylogenetica is BEAST2 een van de meest gebruikte hulpprogramma's. 
    Het is gebundeld met de grafische gebruiksinterface BEAUti 2, DensiTree en Tracer, 
    om BEAST2-configuratiebestanden te maken en om BEAST2-outputbestanden te interpreteren. 
    Echter, als veel verschillende aligneringen of modelopzetten nodig zijn, 
    is een werkvolgorde van meerdere grafische gebruiksinterfaces onhandig. \\
  \textbf{2. }
    Hier presenteren we een gratis, vrij en open-source package, \verb;babette;: 
    'BEAUti 2, BEAST2 en Tracer voor R', voor de programmeertaal R. 
    \verb;babette; schrijft BEAST2-configuratiebestanden, start BEAST2 and verwerkt de resultaten, 
    alles met een enkele R functie-aanroep. \\
  \textbf{3. }
    We beschrijven hoe \verb;babette; te gebruiken is 
    en de nieuwe mogelijkheden die het biedt vergeleken met de originele programma's, 
    aan de hand van enkele voorbeelden. \\
  \textbf{4. }
    Omdat \verb;babette; ontworpen is voor uitbreidbaarheid en hoge kwaliteit, 
    sluiten we af met het beschrijven van de verdere ontwikkeling van dit package. \\
\end{abstract}

@@

However, BEAUti cannot be called from a command-line script.
This implies that when the user 
wants to explore the consequences of various settings, this must be done manually.
This is the managable workflow when using a few alignments and doing a superficial 
analysis of sensitivity of the reconstructed tree to model settings. 
For exploring many trees (for instance from simulations), 
for a sliding-window analysis on a genomic alignment, 
or for a more thorough sensitivity analysis, one would like to loop through 
multiple (simulated or shortened) alignments, nucleotide substitution models, 
clock models and tree priors. 
One such tool to replace BEAUti is \verb;BEASTmasteR; (\cite{beastmaster}),
which focuses on morphological traits and tip-dating, but also 
supports DNA data. \verb;BEASTmasteR;, however, requires hundreds of
lines of R code to setup the BEAST2 model configuration 
and a Microsoft Excel file to specify alignment files.
